Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-18 acknowledgements/acknowledgements.tex
Updates from Overleaf

I want to express my sincerest gratitude to...

\begin{itemize}
\item Carlos Lordelo, and Carl Thomé for their invaluable guidance and support throughout my thesis. Your expertise in MIR, ML, and DSP helped shape the direction of my research and make it a success. Your unwavering support and encouragement helped me overcome any difficulties I encountered. I also appreciate the time and effort you invested in mentoring me. Your dedication to your students is truly inspiring, and I am honored to have had the opportunity to work with you.
\vspace*{3mm}
\item Cody Hesse and Sebastian Löf
\vspace*{3mm}
\item Epidemic Sound AB (\textit{Aktiebolag})
\vspace*{3mm}
\item My co-supervisor
\vspace*{3mm}
\item The Music Technology Group (MTG)
\vspace*{3mm}
\item My colleague at My Sheet Music Transcriptions, Oriol López, for his exceptional support and flexibility throughout my thesis. His understanding and willingness to accommodate my academic commitments effectively allowed me to balance my work and research responsibilities. His guidance and advice on professional matters have been instrumental in my growth as a professional. I am genuinely grateful for his unwavering support and encouragement and for being a great mentor.
\vspace*{3mm}
Thank you all for being excellent advisors and playing a crucial role in completing my thesis. I sincerely appreciate everything you have done for me and will always cherish the knowledge and experience I have gained under your guidance.
\vspace*{3mm}
\item My family
\vspace*{3mm}
\item My girlfriend
\vspace*{3mm}
\end{itemize}

>>>>

I want to extend my deepest gratitude to...

\begin{itemize}
\item Carlos Lordelo and Carl Thomé for their invaluable guidance and support throughout my thesis. Their MIR, ML, and DSP expertise has been instrumental in shaping my research and driving it toward success. I am truly grateful for their unwavering support and encouragement, which helped me surmount any difficulties I encountered during this journey. Their time, effort, and dedication in mentoring me have not gone unnoticed. They inspire me with their commitment to their students, and I feel honored to have had the opportunity to work with them.

Words fall short of my profound gratitude for their consistent, knowledgeable, and steadfast support. I am beyond grateful that their dedication has left an indelible mark on my academic journey.

\vspace*{3mm}
\item Cody Hesse and Sebastian Löf
\vspace*{3mm}
\item Epidemic Sound AB (\textit{Aktiebolag})
\vspace*{3mm}
\item My co-supervisor
\vspace*{3mm}
\item The Music Technology Group (MTG)
\vspace*{3mm}
\item My colleague at My Sheet Music Transcriptions, Oriol López, for his exceptional support and flexibility throughout my thesis. His understanding and willingness to accommodate my academic commitments effectively allowed me to balance my work and research responsibilities. His guidance and advice on professional matters have been instrumental in my growth as a professional. I am genuinely grateful for his unwavering support and encouragement and for being a great mentor.
\vspace*{3mm}
Thank you all for being excellent advisors and playing a crucial role in completing my thesis. I sincerely appreciate everything you have done for me and will always cherish the knowledge and experience I have gained under your guidance.
\vspace*{3mm}
\end{itemize}

### 2023-05-18 dedication/dedication.tex
Updates from Overleaf

%\addcontentsline{toc}{chapter}{Acknowledgement}
\begin{dedication}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
I dedicate this master's thesis to my closest family (Joaquim Colomé, Imma Font, Maria Colomé, and Ariadna Balasch), who has been my constant support and inspiration throughout this journey. Their unwavering love and encouragement have motivated me to pursue my academic goals, and their belief in me has never faltered. I am grateful for their patience, understanding, and sacrifices as I pursued my studies. Their presence in my life has made this journey meaningful, and this thesis is a tribute to their unwavering support. 
\newpage

>>>>

%\addcontentsline{toc}{chapter}{Acknowledgement}
\begin{dedication}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
I dedicate this master's thesis to my closest family and friends, who have supported and inspired me throughout this journey. Their unwavering love and encouragement have motivated me to pursue my academic goals, and their belief in me has never faltered. I am grateful for their patience, understanding, and sacrifices as I pursued my studies.

Their presence in my life has made this journey meaningful, and this thesis is a tribute to their support.

\newpage

### 2023-05-18 discussion/discussion.tex
Updates from Overleaf

>>>>

In this work, we introduced a method to uncover underlying high-level musical
content in the temporal domain by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music latent representations with applications to MIR downstream tasks. Building on top of existing approaches and architectures and wrapped within MSAF \cite{NietoMSAF:FRAMEWORK}, we replaced traditional features with deep embeddings trained to represent high-level musical content information.

### 2023-05-18 evaluation/subsections/beyond dsp.tex
Updates from Overleaf

\section{SALAMI dataset}
\subsection{Overview}

@@

\subsubsection{Formal Analysis Definition}
Formal analysis, in this context, doesn't refer to classifying the piece into a formal type (e.g., sonata, song, or canon) or reducing it into its Ursatz as in Schenkerian analysis. Instead, it means organizing and dividing the piece into specific sections and understanding how they relate.

%%%%%%%%%%

\subsubsection{Approaches to Music Analysis}

The method proposed combines aspects of the following three approaches, separating the organization of instrumentation, musical material, and formal function. This allows for the analysis of a wide variety of music pieces in a consistent way using a constrained vocabulary. It's important to note that music structure is often hierarchical, and the appropriate timescale for analysis can be hard to determine. The proposed method includes a few markers to partly address this issue.

Perceptual Approach: This involves listening for prominent harmonic or rhythmic boundaries in the piece to segment and apply labels to sections based on their similarity. However, this approach might not reflect the different structural roles of similar-sounding sections.

Functional Approach: This involves dividing the song into its functional parts, like verses, choruses, intros, outros, etc. This might result in the same music being labeled differently based on its function or different musical ideas having the same functional label.

Transcription Approach: This involves recording musical parameters like chord patterns and instrumentation directly. While this method may be less subjective, it may not provide a comprehensive view of the piece's structure.

\section{\textit{Embeddiogram} custom feature}

Features extracted from audio files play a vital role in characterizing and differentiating sound and music. One such feature, the \textit{Embeddiogram}, offers a unique and powerful representation of an audio signal. The Embeddiogram is derived by applying a pre-trained neural network model to windowed segments of an audio signal, generating a sequence of embedding vectors. These embedding vectors collectively form a high-dimensional description of the audio signal's structure and content.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/embeddiogram_SALAMI_track_2.png}
    \caption[Embeddiogram]{Caption}
    \label{fig:embeddiogram}
\end{figure}

>>>>

\subsection{SALAMI dataset}

@@

\subsubsection{Formal analysis definition and approaches}
Formal analysis, in this context, doesn't refer to classifying the piece into a formal type (e.g., sonata, song, or canon) or reducing it into its Ursatz as in Schenkerian analysis. Instead, it means organizing and dividing the piece into specific sections and understanding how they relate.

The proposed method integrates aspects of three approaches: perceptual, functional, and transcription, enabling a comprehensive analysis of a wide range of music pieces using a limited vocabulary.

It considers the hierarchical nature of music structure and employs markers to partially address the challenge of determining the appropriate timescale for analysis.

The perceptual approach involves segmenting and labeling sections based on harmonic or rhythmic boundaries, although it may not accurately reflect the structural roles of similar-sounding sections.

The functional approach categorizes the song into components like verses and choruses, potentially leading to varied labels for identical music or identical labels for different musical ideas based on their function.

The transcription approach records musical parameters directly, such as chord patterns and instrumentation, providing a less subjective method, but it might not capture the complete structure of the piece.

\section{\textit{Embeddiogram} feature}

The Embeddiogram is derived by applying a pre-trained neural network model to windowed segments of an audio signal, generating a sequence of embedding vectors. These embedding vectors collectively form a high-dimensional description of the audio signal's structure and content.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/embeddiogram_SALAMI_track_2.png}
    \caption[Embeddiogram]{Embeddiogram computed from...}
    \label{fig:embeddiogram}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/noveltypeaks_SALAMI_track_2.png}
    \caption[Novelty curve]{Caption}
    \label{fig:embeddiogram}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/Rec_matrix_SALAMI_track_2.png} % first figure itself
        \caption{first figure}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/Rec_matrix_smooth_SALAMI_track_2.png} % second figure itself
        \caption{second figure}
    \end{minipage}
\end{figure}

### 2023-05-18 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

This section delves into the systematic process of developing, training and evaluating algorithms, which empower data-driven learning, predictions, and decisions. This process encompasses selecting an appropriate algorithm, designing the model architecture, preprocessing and transforming data, and extracting relevant features. By iteratively adjusting the model's parameters, we aim to minimize the prediction error. The choice of techniques, such as supervised, unsupervised, or reinforcement learning, is dictated by the specific problem and data structure. Model evaluation, fine-tuning, and deployment are essential to ensure the model generalizes well to new data and performs effectively in real-world applications.

\subsection{Model architecture and training strategy}

I advocate for exposure as the primary method of learning music, emphasizing the importance of immersing oneself in various musical experiences to foster a holistic understanding of music as an art form. Individuals can cultivate well-rounded musical expertise by engaging with different styles, genres, techniques, and learning approaches, ultimately promoting creativity and experimentation.

Given the nature of our task, we propose using a Triple Siamese Network. This model architecture has been proven efficient for music similarity retrieval tasks \cite{contentmusicsimtriplet2020}. Specifically, we aim to minimize the loss function between an anchor, a positive, and a negative sample by using online triplet mining \cite{Sikaroudi2020OfflinePatches}.

Using the proposed triple siamese network and online triplet mining, we aim to train a model that can effectively distinguish between similar and dissimilar samples while considering their underlying composition and production texture.

\subsection{Self-Supervised Learning (SSL)}

@@

Triplet Siamese Networks offer several advantages, including learning fine-grained similarity metrics, being well-suited for one-shot learning tasks, and offering efficient memory usage and computational complexity due to shared weights. These networks have been applied to various tasks and domains, such as image recognition, face recognition, speaker recognition, text similarity, and medical imaging for diagnosis or treatment purposes.

\subsection{Convolutional Neural Network (CNN)}

A Convolutional Neural Network (CNN) is a deep-learning neural network. The architecture of a CNN is composed of several layers, including:

\begin{itemize}

\item \textbf{Convolutional layers:} These layers apply a convolution operation to the input data. The convolutional layers are designed to effectively learn local patterns and features in the audio data \footnote{For a visual explanation and further understanding of the convolution operation, please refer to this URL: \url{https://youtu.be/KuXjwB4LzSA}}.

Let's consider a 1D input sequence or signal represented as a vector $I = [i_{i}]{n}$ and a filter $K = [k_{j}]_{p}$, where $n$ is the length of the input sequence and $p$ is the length of the filter. The convolution operation is defined as:

\begin{equation}
(I * K){k} = \sum{j=1}^{p} i_{k+j-1}k_{j}
\end{equation}

Here, $k$ ranges from $1$ to $n-p+1$. The resulting feature vector after applying the convolution is denoted by $(I * K)_{k}$.

The convolution operation slides the kernel $K$ over the input sequence $I$, performing an element-wise multiplication and sum of the local region centered at element $k$ to compute the value of each element in the output feature vector, $(I * K)_{k}$. This process is repeated for each component $k$, resulting in a new feature vector that captures local patterns in the input sequence.

In simple terms, the filter or kernel is slid across the 1D input data. For each position, we perform an element-wise multiplication of the filter with the portion of the input it currently covers and sums the results. This value becomes the corresponding entry in the output. The process is repeated for each position the filter can occupy.

\item \textbf{Pooling layers:} These layers downsample the data, reducing the input dimensions while retaining important information. A joint pooling operation is the max-pooling, which can be defined as follows: given an input matrix $I$ and a pooling window of size $m \times n$, the max-pooling operation selects the maximum value within each non-overlapping $m \times n$ region of the input matrix. This operation effectively reduces the input's dimensions while retaining the essential information in each area.
\vspace*{3mm}

\item \textbf{Fully connected layers:} These layers connect every neuron in one layer to every neuron in another, allowing the network to learn non-linear combinations of the features discovered in the previous layers. Mathematically, a fully connected layer can be represented as $Y = XW + b$, where $X$ is the input matrix, $W$ is the weight matrix, $b$ is the bias vector, and $Y$ is the output matrix.
\vspace*{3mm}

\item \textbf{Output layer:} The output layer produces the final predictions of the network.

\end{itemize}

Compared to a traditional neural network, CNNs are more computationally efficient and have less number of parameters to train. This makes them more feasible for large-scale datasets and real-world problems. On top of that, one of the main advantages of using CNNs for audio analysis is that they can automatically and adaptively learn temporal hierarchies of features from audio data, which traditional audio processing methods may not do effectively.

The SampleCNN model \cite{Lee2018SampleCNN:Classification} takes in 1D input data with a single channel. The first layer is a 1D convolutional layer with a kernel size of 3, a stride of 3, and 128 output channels. It is followed by batch normalization and ReLU activation.

After the initial layer, there are nine hidden layers with varying kernel sizes, strides, and channels. Each hidden layer consists of the following:

\begin{itemize}
    \item A 1D convolutional layer with the specified number of input and output channels, kernel size, a stride of 1, and padding of 1.

    \begin{equation}
y_{c_o}(n) = \sum_{c_i=1}^{C_{\text{in}}} \sum_{k=-\lfloor K/2 \rfloor}^{\lfloor K/2 \rfloor} x_{c_i}(n - k) \cdot w_{c_o, c_i}(k)
\end{equation}

Where:

$y_{c_o}(n)$ is the output for the $c_o$-th output channel at position $n$, $C_{\text{in}}$ is the number of input channels, $x_{c_i}(n)$ is the input for the $c_i$-th input channel at position $n$, $w_{c_o, c_i}(k)$ is the kernel weight for the $c_o$-th output channel and $c_i$-th input, channel at position $k$, and $K$ is the kernel size. The stride is set to 1 and padding to 1 (equal padding $\lfloor K/2 \rfloor$ on both sides).

\item Batch normalization. We process a batch of data containing anchor, positive, and negative samples. We pad the waveforms to the same  maximum length among the collection, perform online triplet mining to find the hardest negatives and return the padded anchors, positives, and hardest negatives.

We have intentionally chosen to use zero padding instead of array truncating in our experiment. This decision was based on our belief that preserving the integrity of the music data was more critical, despite the potential trade-off with computational efficiency during the normalization process.

Let $A$, $P$, and $N$ represent the anchor, positive, and negative samples. The maximum array length in the batch is given by:

\item ReLU activation. The Rectified Linear Unit (ReLU) activation function is a simple yet highly effective non-linear function used in neural networks. It helps introduce non-linearity to the model and accelerates training due to its computational efficiency.

\item Max pooling with kernel size and stride equal to the stride value of the current layer.
\end{itemize}

After the hidden layers, there's another 1D convolutional layer with 512 input channels, 512 output channels, a kernel size of 3, a stride of 1, and padding of 1. This is followed by batch normalization and ReLU activation.

If the model is "supervised", dropout with a rate of 0.5 is applied; a regularization technique randomly drops out neurons during training to prevent overfitting, with a 0.5 rate meaning half of the neurons are dropped out.

Finally, a linear layer with 512 input features and 128 output dimension features is applied to generate the embeddings.

@@

\subsubsection{Positive sample generation}
Given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of the chain.

@@

>>>>

This section outlines the methodical procedure of creating, training and assessing algorithms that enable data-driven learning and decision-making. The process involves algorithm selection, model architecture design, data preprocessing and transformation, and feature extraction. The objective is to minimize prediction error by iteratively fine-tuning the model's parameters. The selection between supervised, unsupervised, or reinforcement learning techniques depends on the problem's specifics and data structure. Evaluating, refining, and deploying the model to ensure its effective performance with new data in real-world applications is crucial.

I advocate for exposure as the primary method of learning music, highlighting the importance of various musical experiences for a comprehensive understanding of the art form. Engagement with diverse musical styles, genres, techniques, and learning methods fosters well-rounded musical proficiency, encouraging creativity and experimentation.

Considering the task at hand, we suggest using a Triple Siamese Network, an architecture proven efficient for music similarity retrieval tasks \cite{contentmusicsimtriplet2020}. The goal is to minimize the loss function between an anchor, positive, and negative sample, achieved through online triplet mining \cite{Sikaroudi2020OfflinePatches}.

Using the proposed triple siamese network, and online triplet mining in a self-supervised setting, we aim to train a model that can effectively distinguish between similar and dissimilar samples while considering their underlying composition and production texture.

The SampleCNN model \cite{Lee2018SampleCNN:Classification} takes 1D input data with a single channel. The first layer is a 1D convolutional layer with a kernel size of 3, a stride of 3, and 128 output channels. It is followed by batch normalization and ReLU activation.

%%%%%BATCH NORM%%%%%%%%%%

The maximum array length in the batch is given by:

Due to shared weights, Triplet Siamese Networks offer efficient memory usage and computational complexity.

\subsubsection{Self-Supervised Learning (SSL)}

@@

@@

\subsubsection{Positive sample generation}

While useful audio transformations tools \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0}

Given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of the chain.

@@

\subsubsection{Triplet mining}
